---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Survey table: formatted survey estimates

<!-- badges: start -->
<!-- badges: end -->

The `surveytable` package provides short and understandable commands that generate tabulated, formatted, and rounded survey estimates. One useful function, which operates on categorical and logical variables, tabulates estimated counts and percentages with their standard errors and confidence intervals. Other functions list the variables in a survey, estimate the total population, tabulate survey subsets and variable interactions, tabulate numeric variables, tabulate rates, create or modify survey variables, perform t-tests, and save the output. All of the tabulation functions check the NCHS presentation standards to flag low-precision estimates. If the `surveytable` code is called from an R Markdown notebook or a Quarto document, it generates HTML tables, which can be incorporated directly into documents. 

## Installation

You can install `surveytable` like so:

``` r
install.packages(c("remotes", "git2r"))
remotes::install_github("CDCgov/surveytable", upgrade = "never")
```

Once `surveytable` is available on CRAN, you will be able to install it like this:

``` r
install.packages("surveytable")
```

## Documentation

Find the documentation for `surveytable` here: https://cdcgov.github.io/surveytable/


## Example

Here is a basic example, to get you started. 

```{r, results=FALSE, message=FALSE, warning=FALSE}
library(surveytable)
```

```{r}
set_survey("namcs2019sv")
```

```{r, results='asis'}
tab("AGER")
```

<!-- CDC standard text -->
## Public Domain Standard Notice
This repository constitutes a work of the United States Government and is not
subject to domestic copyright protection under 17 USC ยง 105. This repository is in
the public domain within the United States, and copyright and related rights in
the work worldwide are waived through the [CC0 1.0 Universal public domain dedication](https://creativecommons.org/publicdomain/zero/1.0/).
All contributions to this repository will be released under the CC0 dedication. By
submitting a pull request you are agreeing to comply with this waiver of
copyright interest.

## License Standard Notice
The repository utilizes code licensed under the terms of the Apache Software
License and therefore is licensed under ASL v2 or later.

This source code in this repository is free: you can redistribute it and/or modify it under
the terms of the Apache Software License version 2, or (at your option) any
later version.

This source code in this repository is distributed in the hope that it will be useful, but WITHOUT ANY
WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A
PARTICULAR PURPOSE. See the Apache Software License for more details.

You should have received a copy of the Apache Software License along with this
program. If not, see https://www.apache.org/licenses/LICENSE-2.0.html

The source code forked from other open source projects will inherit its license.

## Privacy Standard Notice
This repository contains only non-sensitive, publicly available data and
information. All material and community participation is covered by the
[Disclaimer](https://github.com/CDCgov/template/blob/master/DISCLAIMER.md)
and [Code of Conduct](https://github.com/CDCgov/template/blob/master/code-of-conduct.md).
For more information about CDC's privacy policy, please visit [https://www.cdc.gov/other/privacy.html](https://www.cdc.gov/other/privacy.html).

## Contributing Standard Notice
Anyone is encouraged to contribute to the repository by [forking](https://docs.github.com/articles/fork-a-repo)
and submitting a pull request. (If you are new to GitHub, you might start with a
[basic tutorial](https://docs.github.com/articles/set-up-git).) By contributing
to this project, you grant a world-wide, royalty-free, perpetual, irrevocable,
non-exclusive, transferable license to all users under the terms of the
[Apache Software License v2](https://www.apache.org/licenses/LICENSE-2.0.html) or
later.

All comments, messages, pull requests, and other submissions received through
CDC including this GitHub page may be subject to applicable federal law, including but not limited to the Federal Records Act, and may be archived. Learn more at [https://www.cdc.gov/other/privacy.html](https://www.cdc.gov/other/privacy.html).

## Records Management Standard Notice
This repository is not a source of government records, but is a copy to increase
collaboration and collaborative potential. All government records will be
published through the [CDC web site](https://www.cdc.gov).

## Additional Standard Notices
Please refer to [CDC's Template Repository](https://github.com/CDCgov/template)
for more information about [contributing to this repository](https://github.com/CDCgov/template/blob/master/CONTRIBUTING.md),
[public domain notices and disclaimers](https://github.com/CDCgov/template/blob/master/DISCLAIMER.md),
and [code of conduct](https://github.com/CDCgov/template/blob/master/code-of-conduct.md).
