---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Survey table: streamlined output from the `survey` package

<!-- badges: start -->
<!-- badges: end -->

In R, the standard way of analyzing complex surveys is using the `survey` package. The major purpose of the `surveytable` package is to ease the use of the `survey` package in certain applications. `surveytable` improves the output of `survey` functions by formatting and tabulating it; performs hypothesis tests; reduces the number of commands that users need to execute; and applies presentation standards to estimates, which is the usual practice at the National Center for Health Statistics (NCHS). 

The `surveytable` package allows users to use simple commands and produces formatted tabulated output. One useful function, which operates on categorical and logical variables, tabulates estimated counts and percentages with their standard errors and confidence intervals. It applies presentation standards for counts and percentages, and flags estimates if they should be suppressed, footnoted, or reviewed. Other functions list the variables in a survey, estimate the total population, tabulate survey subsets and variable interactions, tabulate numeric variables, calculate rates, perform t-tests, and save the output. 

The `surveytable` package is easier to use than using `survey` directly. With fewer commands, `surveytable` output has survey name, variable labels, formatted estimates, cleaner category values, both count and percentage estimates in a single table, and confidence intervals. The package also performs checking for presentation standards. 

## Installation

You can install `surveytable` like so:

``` r
install.packages("remotes")
remotes::install_git(
	url = "https://github.com/CDCgov/surveytable"
	, upgrade = "never")

```

## Documentation

Find the documentation for `surveytable` here: https://cdcgov.github.io/surveytable


## Example

Here is a basic example, to get you started. 

```{r, results=FALSE, message=FALSE, warning=FALSE}
library(surveytable)
```

```{r, results='asis'}
set_survey("vars2019")
```

```{r, results='asis'}
tab("AGER")
```

## Related packages

In all, there are 3 related packages:

* `surveytable`: functions for tabulating survey estimates
* `nchsdata`: public use files (PUFs) from the the National Center for Health Statistics (NCHS)
* `importsurvey`: functions for importing data into R

`importsurvey` imports surveys into R that are currently in other formats. The public versions of surveys that have already been imported are stored in `nchsdata`. 

<!-- CDC standard text -->
## Public Domain Standard Notice
This repository constitutes a work of the United States Government and is not
subject to domestic copyright protection under 17 USC ยง 105. This repository is in
the public domain within the United States, and copyright and related rights in
the work worldwide are waived through the [CC0 1.0 Universal public domain dedication](https://creativecommons.org/publicdomain/zero/1.0/).
All contributions to this repository will be released under the CC0 dedication. By
submitting a pull request you are agreeing to comply with this waiver of
copyright interest.

## License Standard Notice
The repository utilizes code licensed under the terms of the Apache Software
License and therefore is licensed under ASL v2 or later.

This source code in this repository is free: you can redistribute it and/or modify it under
the terms of the Apache Software License version 2, or (at your option) any
later version.

This source code in this repository is distributed in the hope that it will be useful, but WITHOUT ANY
WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A
PARTICULAR PURPOSE. See the Apache Software License for more details.

You should have received a copy of the Apache Software License along with this
program. If not, see http://www.apache.org/licenses/LICENSE-2.0.html

The source code forked from other open source projects will inherit its license.

## Privacy Standard Notice
This repository contains only non-sensitive, publicly available data and
information. All material and community participation is covered by the
[Disclaimer](https://github.com/CDCgov/template/blob/master/DISCLAIMER.md)
and [Code of Conduct](https://github.com/CDCgov/template/blob/master/code-of-conduct.md).
For more information about CDC's privacy policy, please visit [http://www.cdc.gov/other/privacy.html](https://www.cdc.gov/other/privacy.html).

## Contributing Standard Notice
Anyone is encouraged to contribute to the repository by [forking](https://help.github.com/articles/fork-a-repo)
and submitting a pull request. (If you are new to GitHub, you might start with a
[basic tutorial](https://help.github.com/articles/set-up-git).) By contributing
to this project, you grant a world-wide, royalty-free, perpetual, irrevocable,
non-exclusive, transferable license to all users under the terms of the
[Apache Software License v2](http://www.apache.org/licenses/LICENSE-2.0.html) or
later.

All comments, messages, pull requests, and other submissions received through
CDC including this GitHub page may be subject to applicable federal law, including but not limited to the Federal Records Act, and may be archived. Learn more at [http://www.cdc.gov/other/privacy.html](http://www.cdc.gov/other/privacy.html).

## Records Management Standard Notice
This repository is not a source of government records, but is a copy to increase
collaboration and collaborative potential. All government records will be
published through the [CDC web site](http://www.cdc.gov).

## Additional Standard Notices
Please refer to [CDC's Template Repository](https://github.com/CDCgov/template)
for more information about [contributing to this repository](https://github.com/CDCgov/template/blob/master/CONTRIBUTING.md),
[public domain notices and disclaimers](https://github.com/CDCgov/template/blob/master/DISCLAIMER.md),
and [code of conduct](https://github.com/CDCgov/template/blob/master/code-of-conduct.md).
