---
title: "Printing (HTML)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Printing (HTML)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Basic printing

If a tabulation function is called from the top level, it should print out its table(s) on its own. 

As usual, first, let's start up the package and pick a survey to analyze:

```{r, results='asis'}
library(surveytable)
set_survey(namcs2019sv)
```

Now, when a tabulation function is called from the top level, it prints. You don't need to do anything extra.

```{r, results='asis'}
tab("AGER")
```

If that tabulation function is called not from the top level, such as from within a loop or another function, you need to call `print()` explicitly for it to print. For example:

```{r, results='asis'}
for (vr in c("AGER", "SEX")) {
  print( tab_subset(vr, "MAJOR", "Preventive care") )
}
```

# Create HTML or LaTeX tables

To create HTML or LaTeX tables from an R Markdown notebook or a Quarto document, add the `results='asis'` argument to your code chunk, like so:

<!-- https://yihui.org/knitr/faq/ --> 

````{verbatim}
```{r, results='asis'}
tab("AGER")
```
````

The above should produce the following:

```{r, results='asis', echo=FALSE}
tab("AGER")
```

# Print using various table-making packages

You can change the package that `surveytable` uses for printing. `surveytable` comes with code for using one of these packages: `huxtable` (default), `gt`, and `kableExtra`. In addition, you can supply custom code to use another table-making package.

Changing the table-making package has a couple of uses:

* Use `as_object()` to generate an object from your favorite table-making package, further edit this object using that table-making package, and then finally print it, so the table looks exactly the way you want it to look.
* Print to destinations other than the screen, such as HTML or LaTeX.

## `huxtable`

By default, `surveytable` prints using `huxtable`. If you've changed this, you can always go back with: 

```{r}
set_opts(output = "huxtable")
```

This is what printing to the screen looks like. (To make the table slightly narrower, to make sure that it fits on the page, we temporarily round counts to the nearest thousand.)

```r
set_opts(count = "1k")
tab("AGER")
set_opts(count = "int")
```
```{r, echo=FALSE}
set_opts(count = "1k")
print(tab("AGER"), destination = "")
set_opts(count = "int")
```

To create HTML or LaTeX tables from an R Markdown notebook or a Quarto document, add the `results='asis'` argument to your code chunk, like so:

````{verbatim}
```{r, results='asis'}
tab("AGER")
```
````

This should produce the following:

```{r, results='asis', echo=FALSE}
tab("AGER")
```

## `gt`

```{r}
set_opts(output = "gt")
```

With `gt`, printing to the screen and to HTML look the same. Here is what printing to the screen looks like:

```r
tab("AGER")
```
```{r, results='asis', echo=FALSE}
print(tab("AGER"), destination = "")
```

Here is HTML:

````{verbatim}
```{r, results='asis'}
tab("AGER")
```
````

This should produce the following:

```{r, results='asis', echo=FALSE}
tab("AGER")
```


## `kableExtra`

```{r}
set_opts(output = "kableExtra")
```

We have not implemented screen printing with `kableExtra` yet. Try one of the other packages.

Here is HTML:

````{verbatim}
```{r, results='asis'}
tab("AGER")
```
````

This should produce the following:

```{r, results='asis', echo=FALSE}
tab("AGER")
```

# Advanced printing

## The proper approach

Advanced users can add functionality to use **any** table-making package that they want. For more information, see `help("surveytable-options")`.

## The "quick-and-dirty" approach

The tabulation functions return either:

* for a single table, a data frame, with certain attributes set; or
* for more than one table, a list of such data frames. 

You can convert a single table to a `data.frame` with `as.data.frame()`, like so:

```{r}
tab("AGER") |> as.data.frame()
```

Alternatively, you can pass this data frame to your favorite table-making package. This example passes it to `gt`:

```{r}
set_opts(count = "1k")
tab("AGER") |> gt::gt()
```

The reason that this is the "quick-and-dirty" approach is that the output it creates is not as nice as conventional tables, described above. The output does not have table title (which has important information about the variable and the survey), table footer (which has important information about sample size and low-precision estimates), and it does not format the estimates. Nevertheless, there could be situations in which this approach is helpful, such as 

* extracting an exact value from a table using `as.data.frame()`; or 
* quickly using your favorite table-making package.

