---
title: "Printing (HTML)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Printing (HTML)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Basic printing

If a tabulation function is called from the top level, it should print out its table(s) on its own. 

As usual, first, let's start up the package and pick a survey to analyze:

```{r, results='asis'}
library(surveytable)
set_survey(namcs2019sv)
```

Now, when a tabulation function is called from the top level, it prints. You don't need to do anything extra.

```{r, results='asis'}
tab("AGER")
```

If that tabulation function is called not from the top level, such as from within a loop or another function, you need to call `print()` explicitly for it to print. For example:

```{r, results='asis'}
for (vr in c("AGER", "SEX")) {
  print( tab_subset(vr, "MAJOR", "Preventive care") )
}
```

# Create HTML or LaTeX tables

To create HTML or LaTeX tables from an R Markdown notebook or a Quarto document, add the `results='asis'` argument to your code chunk, like so:

<!-- https://yihui.org/knitr/faq/ --> 

````{verbatim}
```{r, results='asis'}
tab("AGER")
```
````

The above should produce the following:

```{r, results='asis', echo=FALSE}
tab("AGER")
```

# Print using various printing packages

You can change the package that `surveytable` uses for printing. `surveytable` comes with code for using one of these packages: `huxtable` (default), `gt`, and `kableExtra`. In addition, you can supply custom code to use another printing package.

Changing the printing package has a couple of uses:

* Use `as_object()` to generate an object from your favorite printing package, further edit this object using that printing package, and then finally print it, so the table looks exactly the way you want it to look.
* Print to destinations other than the screen, such as HTML or LaTeX.

## `huxtable`

By default, `surveytable` prints using `huxtable`. If you've changed this, you can always go back with: 

```{r}
set_opts(output = "huxtable")
```

This is what printing to the screen looks like. (To make the table slightly narrower, to make sure that it fits on the page, we temporarily round counts to the nearest thousand.)

```r
set_opts(count = "1k")
tab("AGER")
set_opts(count = "int")
```
```{r, echo=FALSE}
set_opts(count = "1k")
print(tab("AGER"), destination = "")
set_opts(count = "int")
```

To create HTML or LaTeX tables from an R Markdown notebook or a Quarto document, add the `results='asis'` argument to your code chunk, like so:

````{verbatim}
```{r, results='asis'}
tab("AGER")
```
````

This should produce the following:

```{r, results='asis', echo=FALSE}
tab("AGER")
```

## `gt`

```{r}
set_opts(output = "gt")
```

With `gt`, printing to the screen and to HTML look the same. Here is what printing to the screen looks like:

```r
tab("AGER")
```
```{r, results='asis', echo=FALSE}
print(tab("AGER"), destination = "")
```

Here is HTML:

````{verbatim}
```{r, results='asis'}
tab("AGER")
```
````

This should produce the following:

```{r, results='asis', echo=FALSE}
tab("AGER")
```


## `kableExtra`

```{r}
set_opts(output = "kableExtra")
```

We have not implemented screen printing with `kableExtra` yet. Try one of the other packages.

Here is HTML:

````{verbatim}
```{r, results='asis'}
tab("AGER")
```
````

This should produce the following:

```{r, results='asis', echo=FALSE}
tab("AGER")
```
