---
title: "Printing (PDF)"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{Printing (PDF)}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
pkgdown:
  as_is: true
  extension: pdf
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Basic printing

If a tabulation function is called from the top level, it should print out its table(s) on its own. 

As usual, first, let's start up the package and pick a survey to analyze:

```r
library(surveytable)
set_survey(namcs2019sv)
```

```{r, results='asis', echo=FALSE, message=FALSE}
library(surveytable)
set_opts(output = "kableExtra")
set_survey(namcs2019sv)
```

Now, when a tabulation function is called from the top level, it prints. You don't need to do anything extra.

```{r, results='asis'}
tab("AGER")
```

If a tabulation function is called not from the top level, such as from within a loop or another function, you do need to call `print()` explicitly for it to print. For example:

```{r, results='asis'}
for (vr in c("AGER", "SEX")) {
  print( tab_subset(vr, "MAJOR", "Preventive care") )
}
```

# Create HTML or PDF tables

Using a Quarto document, you can create tables in many different formats, such as HTML or PDF. Here is a straightforward example of what a Quarto document might look like: 

<!-- https://yihui.org/knitr/faq/ --> 

````{verbatim}
---
title: "My tables"
author: "Me"
format: pdf
---

# Welcome 

As usual, first, let's start up the package and pick a survey to analyze:

```{r, results='asis'}
library(surveytable)
set_survey(namcs2019sv, output = 'auto')
```

# Tables

Take a look at this table:

```{r, results='asis'}
tab("AGER")
```
````

Note the `format` setting, which specifies that this document will create PDF tables. Also note that you do have to add the `results='asis'` argument to the code chunks that print tables. 

# Print using various table-making packages

Use the `output` argument of `set_opts()` to select a table-making package. By default (`output = "auto"`), `surveytable` automatically selects a package depending on whether the output is to the screen (`huxtable`), HTML (`gt`), or PDF (`kableExtra`). You can also explicitly select one of these packages.

Changing the table-making package has a couple of uses:

* It allows you to generate tables in the way that you prefer.
* It allows you to print those tables to a variety of destinations, including the screen, HTML files, or PDF files.

## `kableExtra`

```{r}
set_opts(output = "kableExtra")
```

We have not implemented screen printing with `kableExtra` yet. Try one of the other packages.

Here is PDF:

````{verbatim}
```{r, results='asis'}
tab("AGER")
```
````
```{r, results='asis', echo=FALSE}
tab("AGER")
```

## `auto`

`auto` is the default option. It automatically selects one of the above packages depending on whether the output is to the screen (`huxtable`), HTML (`gt`), or PDF (`kableExtra`).

```{r}
set_opts(output = "auto")
```

PDF output (this should use `kableExtra`):

````{verbatim}
```{r, results='asis'}
tab("AGER")
```
````

```{r, results='asis', echo=FALSE}
tab("AGER")
```

# Generate unformatted output

Some analysts might wish to compare the output from `surveytable` to the output from other statistical software, such as SAS / SUDAAN. In this situation, `set_opts(output = "raw")` might be useful. This command tells `surveytable` to print unformatted and unrounded tables.  

```{r}
set_opts(output = "raw")
```
```r
tab("AGER")
```
```{r, echo=FALSE}
print(tab("AGER"), destination = "")
```
```{r}
set_opts(output = "auto")
```

# Save the tables

## Save tables and charts to an Excel workbook

Before using Excel printing, please be sure to install these packages: `openxlsx2` and `mschart`.

To save tables and charts to an Excel file, turn on Excel printing with `set_opts( output = "Excel", file = "my_workbook" )`. Set the `file` argument to the name of an Excel file. 

```r
set_opts(output = "excel", file = "my_workbook")
```

```{r, echo=FALSE}
set_opts(output = "excel", file = "my_workbook", .file_temp = TRUE)
```

Generate some tables:

```{r}
total()
tab("AGER")
```

To turn off Excel printing, set the `output` argument to a value other than `"Excel"`, such as `"auto"`:

```{r}
set_opts(output = "auto")
```

## Save to a CSV file

To save tables to a CSV file, turn on CSV printing with `set_opts( output = "CSV", file = "my_output" )`. Set the `file` argument to the name of a CSV file. 

```r
set_opts(output = "csv", file = "my_output")
```

```{r, echo=FALSE}
set_opts(output = "csv", file = "my_output", .file_temp = TRUE)
```

Generate some tables:

```{r}
total()
tab("AGER")
```

To turn off CSV printing, set the `output` argument to a value other than `"CSV"`, such as `"auto"`:

```{r}
set_opts(output = "auto")
```

## Save to an R data file

Use the built-in `saveRDS()` function to save a table to an R data file:

```r
tab("AGER") |> saveRDS("myfile.rds")
```

You can later load this data file back into R. To print the table, just load the file, like so:

```r
readRDS("myfile.rds")
```

```{r, results='asis', echo=FALSE}
tab("AGER")
```


# Advanced printing

## The proper approach

Advanced users can add functionality to use **any** table-making package that they want. For more information, see `help("surveytable-options")`.

## The "quick-and-dirty" approach

The tabulation functions return either:

* for a single table, which is basically a data frame, with certain attributes set; or
* for more than one table, a list of such tables.

You can convert a single table to a `data.frame` with `as.data.frame()`, like so:

```{r}
tab("AGER") |> as.data.frame()
```

Note that this produces a `data.frame` with unique column names, which improves its usability.

Alternatively, you can pass this `data.frame` to your favorite table-making package. This example passes a table to `gt`. To ensure unique column names, pass the table through `as.data.frame()` first.

```{r}
tab("AGER") |> as.data.frame() |> gt::gt()
```

The reason that this is the "quick-and-dirty" approach is that the output it creates is not as nice as conventional tables, described above. The output does not have table title (which has important information about the variable and the survey), table footer (which has important information about sample size and low-precision estimates), and it does not format the estimates. Nevertheless, there could be situations in which this approach is helpful, such as 

* extracting an exact value from a table using `as.data.frame()`; or 
* quickly using your favorite table-making package.
