---
title: "Introduction to surveytable"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{surveytable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of `surveytable` is to generate streamlined output from the `survey` package, with an application to NCHS.

The examples below use the National Ambulatory Medical Care Survey (NAMCS) 2019 Public Use File (PUF). NAMCS is "an annual nationally representative sample survey of visits to nonfederal office-based patient care physicians, excluding anesthesiologists, radiologists, and pathologists." Note that the unit of observation is visits, not patients – this distinction is important since a single patient can make multiple visits.

Selected variables from NAMCS 2019 come with the `surveytable` package, in an object called `vars2019`. Alternatively, if you have installed the `nchsdata` package, to access the full NAMCS 2019, use `nchsdata::namcs2019`.

To make the below code more reusable, we'll call the survey by a generic name, namely, `mysurvey`.

# Begin

Begin by loading the `surveytable` package. When you do, it will print a message explaining how to specify the survey that you'd like to analyze. We are not showing that message here. 

```{r, results=FALSE, message=FALSE, warning=FALSE}
library(surveytable)
```

Now, specify the survey which you'd like to analyze.

```{r, results='asis'}
mysurvey = vars2019
# Or, if nchsdata has been installed, 
# mysurvey = nchsdata::namcs2019
set_survey("mysurvey")
```

Check the survey name, survey design variables, and the number of observations to verify that it all looks correct.

# List variables

List the variables that start with the letters `age`:

```{r, results='asis'}
var_list("age")
```

The table lists variable name, class (or the type of variable), and variable label, which is the long name of the variable. Common classes are `factor` (categorical variable), `logical` (yes / no variable), and numeric.

# Tabulate categorical and logical variables
```{r, results='asis'}
tab("AGER")
```

The table title shows the variable label (the long variable name) and the survey label.

For each level of the variable, the table shows:

* estimated count, its standard error, and the 95% confidence interval
* estimated percentage, its standard error, and the 95% confidence interval

If there are flags for presentation standards for counts and percentages, they are shown in an additional column and in the footnote. 

Several more tables: 

```{r, results='asis'}
tab("MDDO", "SPECCAT", "MSA")
```

# Entire population
To estimate the total count for the entire population:

```{r, results='asis'}
total()
```

# Tabulate subsets or interactions
```{r, results='asis'}
tab_subset("AGER", "SEX")
```

For each value of `SEX`, the above command creates a table of `AGER`. 

With the `tab_subset` command, the percentages add up to 100% for each value of `SEX`.

The title of each table reflects the value of `SEX`.

Alternatively, you could use this command:

```{r, results='asis'}
tab_cross("AGER", "SEX")
```

This command is similar. The major difference is that with `tab_cross`, percentages add up to 100% for the whole population.

You can also create a new variable by crossing two variables, and use it later:

```{r, results='asis'}
var_cross("Age x Sex", "AGER", "SEX")
var_list("age")
tab("Age x Sex")
```

# Tabulate numeric variables
The above functions also work with numeric variables, though the output is a bit different. 

```{r, results='asis'}
tab("NUMMED")
```

The table shows percent of values that are not missing (not `NA`), mean, standard error of the mean (SEM), and standard deviation (SD). 

Subsetting works too:

```{r, results='asis'}
tab_subset("NUMMED", "AGER")
```

# Perform statistical hypothesis testing
To perform statistical hypothesis testing, use the `test` argument. 

## Categorical variable
```{r, results='asis'}
tab("AGER", test = TRUE)
```

## Categorical variable (subset)
```{r, results='asis'}
tab_subset("AGER", "SEX", test = TRUE)
```

## Numeric variable
```{r, results='asis'}
tab_subset("NUMMED", "AGER", test = TRUE)
```

# Calculate rates
To calculate rates, we need another source of information with population estimates. You would typically use a function like `read.csv()` to load the population estimates and get them into the correct format. The `surveytable` package comes with an object called `uspop2019` that contains several population estimates for use in these examples.

Overall rate:
```{r, results='asis'}
total_rate(uspop2019$total)
```

Rate for a particular variable:
```{r, results='asis'}
tab_rate("AGER", uspop2019$AGER)
```

Rate for an interaction of variables:
```{r, results='asis'}
tab_subset_rate("AGER", "SEX", uspop2019$`AGER x SEX`)
```

# Save the output
Send all future output to a CSV file, create whatever tables you need, then turn off sending output to the file:

```{r, results=FALSE, message=FALSE, warning=FALSE}
set_output(csv = "out.csv")
```

```{r, results='asis'}
tab("MDDO", "SPECCAT", "MSA")
set_output(csv = "")
```

```{r, echo=FALSE}
unlink("out.csv")
```
