---
title: "Introduction to surveytable"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to surveytable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of `surveytable` is to generate streamlined output from the `survey` package, with an application to NCHS.

The examples below use the National Ambulatory Medical Care Survey (NAMCS) 2019 Public Use File (PUF). NAMCS is "an annual nationally representative sample survey of visits to nonfederal office-based patient care physicians, excluding anesthesiologists, radiologists, and pathologists." Note that the unit of observation is visits, not patients – this distinction is important since a single patient can make multiple visits.

Selected variables from NAMCS 2019 come with the `surveytable` package, for use in examples, in an object called `vars2019`. Alternatively, if you have installed the `nchsdata` package, to access the full NAMCS 2019, use `nchsdata::namcs2019`.

# Begin

Begin by loading the `surveytable` package. When you do, it will print a message explaining how to specify the survey that you'd like to analyze. We are not showing that message here. 

```{r, results=FALSE, message=FALSE, warning=FALSE}
library(surveytable)
```

Now, specify the survey that you'd like to analyze.

```{r}
set_survey("vars2019")
```

Check the survey label, survey design variables, and the number of observations to verify that it all looks correct.

## List variables

The `var_list` function lists the variables in the survey. To avoid unintentionally listing all the variables in a survey, which can be many, the starting characters of variable names are specified. For example, to list the variables that start with the letters `age`, type:

```{r, results='asis'}
var_list("age")
```

The table lists 

* variable name; 
* class, which is the type of variable; and 
* variable label, which is the long name of the variable. 

Common classes are `factor` (categorical variable), `logical` (yes / no variable), and `numeric`.

# Tabulate categorical and logical variables

The main function of the `surveytable` package is `tab`, which tabulates variables. It operates on categorical and logical variables, and presents both estimated counts, with their standard errors (SEs) and 95% confidence intervals (CIs), and percentages, with their SEs and CIs. For example, to tabulate `AGER`, a categorical variable of age groups, type:

```{r, results='asis'}
tab("AGER")
```

The table title shows the variable label (the long variable name) and the survey label.

For each level of the variable, the table shows:

* the estimated count, its standard error, and its 95% confidence interval; and 
* the estimated percentage, its standard error, and its 95% confidence interval.

**NCHS presentation standards.** The `tab` function also applies the NCHS presentation standards for counts and percentages, and flags estimates if, according to the standards, they should be suppressed, footnoted, or reviewed by an analyst. The CIs that are displayed are the ones that are used by the NCHS presentation standards. Specifically, for counts, the tables show the log Student's t 95% CI, with adaptations for complex surveys; for percentages, they show the 95% Korn and Graubard CI.

One does not need to do anything extra to perform presentation standards checking. It is performed automatically. For example, let's tabulate `PAYWKCMP` (Expected source of payment for visit: Workers Compensation):

```{r, results='asis'}
tab("PAYWKCMP")
```

This table tells us that, according to the NCHS presentation standards, both the count and the percentage of visits in which the expected source of payment was workers' compensation should be suppressed due to low precision. This makes intuitive sense as the confidence intervals for these estimates are very wide.

**Drop missing values.** Some variables might contain missing values (`NA`). Consider the following variable, which is not part of the actual survey, but was constructed specifically for this example:

```{r, results='asis'}
tab("SPECCAT.bad")
```

To calculate percentages based on the non-missing values only, use the `drop_na` argument:

```{r, results='asis'}
tab("SPECCAT.bad", drop_na = TRUE)
```

The above table gives percentages based only on the knowns, that is, based only on non-`NA` values. 

**Multiple tables.** Multiple tables can be created with a single command: 

```{r, results='asis'}
tab("MDDO", "SPECCAT", "MSA")
```

## Entire population
Estimate the total count for the entire population using the `total` command:

```{r, results='asis'}
total()
```

## Subsets or interactions

To create a table of `AGER` for each value of the variable `SEX`, type:

```{r, results='asis'}
tab_subset("AGER", "SEX")
```

Optionally, one or more levels of the subsetting variable can be specified, to tabulate only on those subsets.

The title of each table reflects the value of the subsetting variable (in this case, `SEX`).

With the `tab_subset` command, in each table (that is, in each subset), the percentages add up to 100%.

Another way to tabulate subsets is using the `tab_cross` function, which crosses or interacts two variables and generates a table using this new variable. While the estimated counts produced by `tab_subset` and `tab_cross` are the same, with `tab_cross`, percentages across the entire population add up to 100%.

```{r, results='asis'}
tab_cross("AGER", "SEX")
```

# Tabulate numeric variables

The `tab` and `tab_subset` functions also work with numeric variables, though with such  variables, the output is different. To tabulate `NUMMED` (number of medications), a numeric variable, type:

```{r, results='asis'}
tab("NUMMED")
```

As before, the table title shows the variable label (the long variable name) and the survey label.

The table shows the percentage of values that are not missing (not `NA`), the mean, the standard error of the mean (SEM), and the standard deviation (SD). 

Subsetting works too:

```{r, results='asis'}
tab_subset("NUMMED", "AGER")
```

# Perform statistical hypothesis testing

The `tab` and `tab_subset` functions make it easy to perform the design-based t-test by using the `test` argument. 

## Categorical variables
To test whether any levels of `AGER` are statistically similar or different, type:

```{r, results='asis'}
tab("AGER", test = TRUE)
```

According to this, the following pairs of categories are statistically similar:

* "25-44 years" and "65-74 years"; and 
* "25-44 years" and "75 years and over".

All other pairs of categories are statistically different. 

## Categorical variables (subset)
To test whether any levels of `AGER` are statistically similar or different within subpopulations defined by the `SEX` variable, type:

```{r, results='asis'}
tab_subset("AGER", "SEX", test = TRUE)
```

According to this, the following pairs of categories are statistically similar:

* Among Females:
  * "25-44 years" and "65-74 years"; and 
  * "25-44 years" and "75 years and over".
* Among Males:
  * "Under 15 years" and "25-44 years";
  * "Under 15 years" and "75 years and over";
  * "25-44 years" and "75 years and over"; and 
  * "65-74 years" and "75 years and over".
  
## Numeric variables

To test whether the numeric variable `NUMMED` varies with the categorical variable `AGER`, type:

```{r, results='asis'}
tab_subset("NUMMED", "AGER", test = TRUE)
```

According to this, `NUMMED` is statistically similar for the "Under 15 years" and "15-24 years" `AGER` categories. It is statistically different for all other pairs of categories. 

# Calculate rates

A rate is a ratio of count estimates based on the survey in question divided by population size, which is assumed to be known. For example, the number of physician visits per 100 people in the population is a rate: the number of physician visits is estimated from the `vars2019` survey, while the number of people in the population comes from another source. 

To calculate rates, in addition to the survey, we need a source of information on population size. You would typically use a function such as `read.csv()` to load the population figures and get them into the correct format. The `surveytable` package comes with an object called `uspop2019` that contains several population figures for use in these examples.

Examine `uspop2019`:

```{r}
class(uspop2019)
names(uspop2019)
```

The overall population size for the country as a whole:
```{r}
uspop2019$total
```

Once we have the overall population size, the overall rate is:
```{r, results='asis'}
total_rate(uspop2019$total)
```

To calculate the rates for a particular variable, we need to provide a data frame with a column called `Level` that matches the levels of the variable in the survey, and a column called `Population` that gives the size of the population for that level.

For example, for `AGER`, this data frame as follows:

```{r}
uspop2019$AGER
```

Now that we have the appropriate population figures, the rates table is obtained by typing:

```{r, results='asis'}
tab_rate("AGER", uspop2019$AGER)
```

To calculate the rates for one variable (`AGER`) by another variable (`SEX`), we need population figures in the following format:
```{r}
uspop2019$`AGER x SEX`
```

With this data frame, the rates table is obtained by typing:

```{r, results='asis'}
tab_subset_rate("AGER", "SEX", uspop2019$`AGER x SEX`)
```

# Create or modify variables

In some situations, it might be necessary to modify survey variables, or to create new ones. This section describes how to do this.

**Collapse levels.** The variable `PRIMCARE` has levels "Unknown" and "Blank", among others. 

```{r, results='asis'}
tab("PRIMCARE")
```

To collapse these two levels into a single level, type:

```{r, results='asis'}
var_collapse("PRIMCARE", "Unknown if PCP", c("Unknown", "Blank"))
tab("PRIMCARE")
```

**Convert numeric to factor.** The variable `AGE` is numeric.

```{r, results='asis'}
tab("AGE")
```

To create a new variable of age categories based on `AGE`, type:

```{r, results='asis'}
var_cut("Age group", "AGE"
        , c(-Inf, 0, 4, 14, 64, Inf)
        , c("Under 1", "1-4", "5-14", "15-64", "65 and over") )
tab("Age group")
```

In the `var_cut` command, specify the following information:

* name of the new categorical variable; 
* name of the existing numeric variable;
* cut points -- note that the intervals are inclusive on the right; and  
* category labels.

**Check whether any variable is true.** For a series of logical variables, you can check whether any of them are `TRUE` using the `var_any` command. 

A physician visit is considered to be an "imaging services" visit if it had any of a number of imaging services ordered or provided. Imaging services are indicated using logical variables, such as `MRI` and `XRAY`. To create the `Imaging services` variable, type:

```{r, results='asis'}
var_any("Imaging services"
  , c("ANYIMAGE", "BONEDENS", "CATSCAN", "ECHOCARD", "OTHULTRA"
  , "MAMMO", "MRI", "XRAY", "OTHIMAGE"))
tab("Imaging services")
```

**Interact variables.** The `tab_cross` function creates a table of an interaction of two variables, but it does not save the interacted variable. An analyst might need to do something with the interacted variable, such as possibly subset it later. To create the interacted variable, use the `var_cross` command:

```{r}
var_cross("Age x Sex", "AGER", "SEX")
```

Now, for example, create a table of the age / sex interaction in preventive care visits:

```{r, results='asis'}
tab_subset("Age x Sex", "MAJOR", "Preventive care")
```

**Copy a variable.** Create a new variable that is a copy of another variable using `var_copy`. You can modify the copy, while the original remains unchanged. For example:

```{r, results='asis'}
var_copy("Age group", "AGER")
var_collapse("Age group", "65+", c("65-74 years", "75 years and over"))
var_collapse("Age group", "25-64", c("25-44 years", "45-64 years"))
tab("AGER", "Age group")
```

**The `variables` data frame.** Recall that survey design objects have an element called `variables`, which is a data frame that contains the survey variables.

```{r}
class(vars2019$variables)
```

More advanced users can create or modify variables in the `variables` data frame directly.

# Save the output

The `tab*` and `total*` functions have an argument called `csv` that specifies the name of a comma-separated values (CSV) file to save the output to. Alternatively, you can name the default CSV output file using the `set_output` function. For example, the following directs `surveytable` to send all future output to a CSV file, create some tables, and then turn off sending output to the file:

```{r, results=FALSE, message=FALSE, warning=FALSE}
set_output(csv = "out.csv")
```

```{r, results='asis'}
tab("MDDO", "SPECCAT", "MSA")
set_output(csv = "")
```

```{r, echo=FALSE}
unlink("out.csv")
```

Tabulation functions also have an argument called `screen`, `TRUE` by default, which determines whether the output is sent to the screen. If many tables are being produced, a user might choose to turn off screen printing, and only output to a CSV file. 

If the tabulation functions are called from within an R Markdown notebook, they produce HTML tables. This makes it easy to incorporate the output of the `surveytable` package directly into documents, presentations, “shiny” web apps, and other output types.

Finally, the tabulation functions return the tables that they produce, either as a data frame, or as a list of data frames. More advanced analysts can use this functionality to integrate `surveytable` into other programming tasks.
