---
title: "Use cases"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use cases}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `surveytable` package can be used with different types of data. Here are 
some examples. 

# Unweighted data

Unweighted data is stored in a `data.frame` or a similar object. One example of such
a similar object is a tibble (`tbl`), from the `tibble` package. These objects do not 
contain information about survey design variables. Thus, `surveytable` treats these
objects as unweighted data, with each observation having a weight of `1`. 

Below, we first create a tibble with unweighted data, and then tabulate the 
`SPECCAT` (physician specialty) variable.

```{r, results='asis'}
library(surveytable)
library(tibble)

mytbl = as_tibble(namcs2019sv_df)

set_survey(mytbl)
tab("SPECCAT")
```

# Complex survey

A survey object, in addition to containing the survey data, contains information
about the survey design variables. These include variables that specify such things as:

* cluster ID’s, also known as primary sampling units (PSU’s);
* cluster sampling probabilities; 
* strata; 
* finite population correction; and 
* sampling weights.

You can convert a `data.frame` or a similar object to a survey object using the 
`survey::svydesign()` command. Before using this command, you should consult the documentation
for the survey that you are analyzing to find out what the survey design variables are.

Below, we first create a survey object, and then tabulate the `SPECCAT` variable.

```{r, results='asis'}
library(surveytable)

mysurvey = survey::svydesign(ids = ~ CPSUM
  , strata = ~ CSTRATM
  , weights = ~ PATWT
  , data = namcs2019sv_df)

set_survey(mysurvey)
tab("SPECCAT")
```

# Complex survey using data stored in Spark

Especially if you are dealing with big data, that data might be stored in a database, 
such as Spark. 

Below, we put some data into a Spark dataset, create a survey object using this Spark dataset, 
and then tabulate the `SPECCAT` variable. Note that, for this example, we are using a 
`"local"` Spark connection -- how you connect to Spark depends on your setup.

```{r, results='asis'}
library(surveytable)
library(sparklyr)
library(dplyr)

sc = spark_connect(master = "local")

mysparkdf = copy_to(sc, namcs2019sv_df)
mysurvey = survey::svydesign(ids = ~CPSUM, strata = ~CSTRATM
                             , weights = ~PATWT, data = mysparkdf)

set_survey(mysurvey)
tab("SPECCAT")

spark_disconnect_all()
```

# Complex survey with replicate weights

Some surveys, instead of survey design variables, specify replicate weights. This
can be done, for example, for privacy reasons.

You can convert a `data.frame` or a similar object to a survey object that uses 
replicate weights using the `survey::svrepdesign()` command.

Below, we first create a survey object using fake replicate and sampling weights, 
and then tabulate the `SPECCAT` variable.

```{r, results='asis'}
library(surveytable)

mydata = namcs2019sv_df
nr = nrow(mydata)
set.seed(42)
for (ii in 1:20) {
  mydata[,paste0("fake_repw", ii)] = runif(nr, 10, 1000)
}
mydata$fake_w = runif(nr, 10, 1000)

mysurvey = survey::svrepdesign(
  repweights = "fake_repw*"
  , weights = ~fake_w
  , data = mydata
)

set_survey(mysurvey)
tab("SPECCAT")
```
